---
title: "Analysis"
author: "Charlotte Dawson"
date: "2018-01-20"
output:
  bookdown::html_document2:
    toc: true
    toc_float: true
    df_print: kable
vignette: > 
  %\VignetteIndexEntry{Vignette title}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r, setup, echo=FALSE, message=FALSE}
knitr::opts_chunk$set(collapse = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      echo = FALSE,
                      comment = "#>")
# use fig.path="../vignettes/figures" to save generated figures
library(limma)
library(dplyr)
library(ggplot2)
library(candidaev)
library(tidyNano)
library(ggpubr)
```

# Introduction

# EV characterisation



```{r, nta_tech_avg}
data(nta)

df <- data.frame(id = c("ATCC10231_1", "ATCC10231_2", "ATCC10231_3", "ATCC90028_1", "ATCC90028_2", "ATCC90028_3", 
                        "Biofilm_1", "Biofilm_2", "Biofilm_3", "Biofilm_4", "Biofilm_5", 
                        "DAY286_1", "DAY286_2", "DAY286_3"), 
                 Protein_conc = c(0.192, 0.282, 0.287, 0.424, 0.716, 0.584, 0.204, 0.198, 0.297, 0.193, 0.416, 0.437, 0.360, 0.447), 
                 Culture_vol = c(rep(150, 6), rep(1000, 5), rep(300, 3)), 
                 OD600 = c(0.455*20, 0.557*20, 0.615*20, 0.310*50, 0.337*50, 0.297*50, rep(NA, 5), 36.2, 39.7, 34.0), 
                 stringsAsFactors = FALSE)

nta_techavg <- nta %>% 
  nanolyze(particle_size, Strain, Bio_rep, Dilution, 
           name = "Tech_rep", 
           param_var = True_count)

nta_techavg %>% 
  ggplot(aes(x = particle_size, y = Tech_rep_mean, colour = Bio_rep)) + 
  geom_line(size = 1) + 
  facet_wrap(~ Strain, scales = "free") + 
  theme_classic()

```

Text.

```{r}
nta_bioavg <- nta_techavg %>% 
  nanolyze(particle_size, Strain,
           name = "Bio_rep", 
           param_var = Tech_rep_mean) %>% 
  mutate(facet = case_when(Strain == "Biofilm" ~ "first", TRUE ~ "second"))

nta_bioavg %>% 
  ggplot(aes(x = particle_size, y = Bio_rep_mean, colour = Strain)) + 
  geom_line(size = 1) + 
  facet_wrap(~ facet, scales = "free") + 
  theme_classic()
  
```

Text

```{r}
nta_total <- nta_techavg %>% 
  nanocount(Strain, Bio_rep, Dilution, 
            param_var = Tech_rep_mean) %>% 
  tidyr::unite("id", Strain, Bio_rep) %>% 
  mutate(protein_conc = as.numeric(match_id(id, df, "id", "Protein_conc")), 
         od = as.numeric(match_id(id, df, "id", "OD600")), 
         part_per_prot = Particle_count / protein_conc / 1000) %>% 
  tidyr::separate(col = id, into = c("Strain", "Bio_rep"), sep = "_", remove = TRUE)

nanoShapiro(nta_total, Strain, value = part_per_prot)

nanoShapiro(nta_total, Strain, value = Particle_count)

```


Text.

```{r}
tukey_hsd <- function(x, ...) {
  res <- TukeyHSD(x, ...) %>%
    broom::tidy() %>%
    mutate(comparison2 = comparison) %>%
    tidyr::separate(comparison2, into= c("group2", "group1"), sep = "-") %>%
    rename(p.adj = adj.p.value) %>%
    mutate(p.adj = signif(p.adj, 2)) %>%
    select(term, group1, group2, everything(), -term)
  res
}

dunns_test <- function(x, g, method, ...) {
  # capture console output before it prints
  dtres <- utils::capture.output(res <- dunn.test::dunn.test(x, g, method, ...))
  
  # convert returned list `res` to data.frame
  res <- res %>% 
    as.data.frame() %>% 
    rename(p.adj = P.adjusted, p.val = P, comparison = comparisons) %>% 
    tidyr::separate(comparison, into = c("group2", "group1"), sep = " - ", remove= FALSE) %>% 
    mutate(p.adj = signif(p.adj, 2)) %>% 
    select(group1, group2, comparison, everything())
    res
}
```


```{r}
# total particle concentration not normally distributed for biofilm
# use Kruskal Wallis then Dunn's test
stat.test <- dunns_test(x = nta_total$Particle_count, 
                        g = nta_total$Strain, method = "bh")

# total particle number per ug protein all normally distributed
# use ANOVA then Tukey's HSD
stat.test2 <- aov(part_per_prot ~ Strain, data = nta_total) %>% 
  tukey_hsd()

# save total particle concentration plot
total_pconc <- nta_total %>% 
  ggboxplot(x = "Strain", y = "Particle_count", 
            color = "Strain", palette = "jco", add = "jitter") 
total_pconc <- ggpar(total_pconc, 
                     yscale = "log10", 
                     format.scale = TRUE, 
                     ylim = c(1e11, 1e13)) + 
  stat_pvalue_manual(stat.test[c(1,2), ], label = "p.adj", 
                     y.position = c(12.4, 13), 
                     tip.length = 4e-3)

# save total particle number per ug protein plot
total_pnum <- nta_total %>% 
  ggboxplot(x = "Strain", y = "part_per_prot", 
            color = "Strain", palette = "jco", add = "jitter")
total_pnum <- ggpar(total_pnum, 
                    yscale = "log10", 
                    format.scale = TRUE, 
                    ylim = c(8e8, 8e10)) + 
  stat_pvalue_manual(stat.test2[c(2,4,6), ], label = "p.adj", 
                     y.position = c(10.7, 10.5, 10.9), 
                     tip.length = 1e-2)

# plot both
ggarrange(total_pconc, total_pnum, 
          labels = c("A", "B"), 
          ncol = 2, 
          nrow = 1, 
          font.label = list(size = 24, color = "black"))
  
```

# Yeast DAY286 

Text.

```{r}

```



# Biofilm DAY286

# Comparison to previous data

# EV marker protein selection

# Different strains test
## Introduction

Aim is to determine whether identified potential biomarkers work in other 
*C. albicans* strains.

## Experimental design info
Strain used:

<div id="left">
- ATCC10231
- zzzzzz
</div>

<div id="right">
- ATCC90028
- zzzzzzz
</div>

```{r, atcc_exp}
atcc_exp <- atcc_exp

```


## Analysis
    
### Prepare data
```{r, load_atcc}
atcc_pg <- atcc

```

The untouched proteinGroups.txt file has `r nrow(atcc_pg)` rows and `r ncol(atcc_pg)`  columns.

Each of the `r nrow(atcc_pg)` rows corresponds to a single protein group. 

```{r, prep_atcc}
atcc <- atcc_pg %>% 
  filter(Reverse != "+", Potential.contaminant != "+", Unique.peptides >= 2)

atcc <- convert_lfq(atcc, atcc_exp)

```

First, filter any reverse and contaminant proteins. 
Then select only proteins identified by at least 2 unique peptides.

Now the dataset contains `r nrow(atcc)` proteins across the 12 samples.

```{r, fig.cap="This is a figure caption", dpi=300}
f7a <- plot_frequency2(atcc)

f7b <- plot_numbers2(atcc, atcc_exp) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))

ggpubr::ggarrange(f7a, f7b,
                  labels = c("A", "B"), 
                  ncol = 2, 
                  nrow = 1, 
                  font.label = list(size = 24, color = "black"))

```


Table goes here.

More text goes here.

```{r}

knitr::kable(summarise_lfq(atcc), 
             caption = "A table caption", 
             col.names = c("Sample", "Mean LFQ intensity", "Median LFQ intensity", 
                           "SD", "CV", "Min LFQ intensity", "Max LFQ intensity", 
                           "Range", "Number of proteins", "Number of NA", "Percent NA"), 
             digits = 1)
```

### Filter data

Filter for proteins identified in 2/3 reps of at least 1 condition.



```{r}
atcc_filt <- filter_na4(atcc, "or", "<=", 
                        pat1 = "10231_EV", 1, 
                        pat2 = "10231_W", 1, 
                        pat3 = "90028_EV", 1, 
                        pat4 = "90028_W", 1)

```

Reduces `r nrow(atcc)` proteins to `r nrow(atcc_filt)`.

```{r, dpi=300, fig.show='hide'}
f8a <- plot_frequency2(atcc_filt)

f8b <- plot_numbers2(atcc_filt, atcc_exp) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))

a1_ev <- filter_na(atcc_filt, "<=", "10231_EV", 2)
a1_w <- filter_na(atcc_filt, "<=", "10231_W", 2)

a9_ev <- filter_na(atcc_filt, "<=", "90028_EV", 2)
a9_w <- filter_na(atcc_filt, "<=", "90028_W", 2)

f8c <- plot_venn(list("10231 ev" = a1_ev, "10231 wcl" = a1_w, 
                      "90028 ev" = a9_ev, "90028 wcl" = a9_w),
                 use_uniprot = TRUE, type = "plot")
f8c <- ggpubr::as_ggplot(f8c) + theme(plot.margin = unit(c(0,5,0,5), "lines"))

```

```{r, dpi=300, fig.height=7}
ggpubr::ggarrange(ggpubr::ggarrange(f8a, f8b, ncol = 2, labels = c("A", "B")), 
                  f8c, 
                  nrow = 2, 
                  labels = c("", "C"))
```


Some words.






































