---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo=FALSE}
knitr::opts_chunk$set(collapse = TRUE,
                      comment = "#>")
```

# vignettes

## Description

This directory contains Rmarkdown files used to generate HTML vignettes that 
detail the proteomic analyses used in our paper: 

> Dawson, C. S., Garcia-Ceron, D., Rajapaksha, H., Faou, P., Anderson, M. A., 
Bleackley, M. R. (2020) Protein markers for _Candida albicans_ EVs include 
claudin-like Sur7 family proteins. _Journal of Extracellular_ 
_Vesicles_, 9(1): 1750810. DOI: 
[10.1080/20013078.2020.1750810](https://doi.org/10.1080/20013078.2020.1750810)

If the `candidaev` package has been installed you can view the vignettes with:

```{r, eval=FALSE}
browseVignettes("candidaev")
```

## Contents
```
    .
    ├── ...
    ├── vignettes
    │    ├── atcc_analysis.Rmd         # Describes ATCC90028 and ATCC10231 EV analyses
    │    ├── biofilm_analysis.Rmd      # Describes DAY286 biofilm EV analysis
    │    └── yeast_analysis.Rmd        # Describes DAY286 yeast EV analysis
    └── ...

```
