---
title: "DAY286 yeast analysis"
author: 
- name: Charlotte Dawson
  affiliation: Department of Biochemistry and Genetics, 
               La Trobe Institute for Molecular Science, 
               La Trobe University, Melbourne, 3086, 
               Victoria, Australia
- name: Donovan Garcia Ceron
  affiliation: Department of Biochemistry and Genetics, 
               La Trobe Institute for Molecular Science, 
               La Trobe University, Melbourne, 3086, 
               Victoria, Australia
- name: Harinda Rajapaksha
  affiliation: La Trobe Comprehensive Proteomics Platform, 
               La Trobe Institute for Molecular Science, 
               La Trobe University, Melbourne, 3086, 
               Victoria, Australia
- name: Pierre Faou
  affiliation: La Trobe Comprehensive Proteomics Platform, 
               La Trobe Institute for Molecular Science, 
               La Trobe University, Melbourne, 3086, 
               Victoria, Australia
- name: Marilyn Anderson
  affiliation: Department of Biochemistry and Genetics, 
               La Trobe Institute for Molecular Science, 
               La Trobe University, Melbourne, 3086, 
               Victoria, Australia
- name: Mark Bleackley
  affiliation: Department of Biochemistry and Genetics, 
               La Trobe Institute for Molecular Science, 
               La Trobe University, Melbourne, 3086, 
               Victoria, Australia
date: "10 April 2020"
output:
  BiocStyle::html_document:
    toc: true
    df_print: kable
package: candidaev
abstract: | 
  Analysis of LC-MS/MS data obtained for _Candida albicans_ (strain = 
  DAY286) extracellular vesicles and yeast whole cell lysate.
vignette: > 
  %\VignetteIndexEntry{DAY286 yeast analysis}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r knitrOpts, include=FALSE}
# knit options
knitr::opts_knit$set(progress = TRUE, 
                     verbose = TRUE)

# chunk options
knitr::opts_chunk$set(collapse = FALSE,
                      warning = TRUE,
                      message = TRUE,
                      echo = FALSE,
                      error = TRUE, 
                      comment = "#>")

# check for installation of packages requires to knit this vignette
req_pkgs <- c("BiocStyle", "ggpubr", "stringi", "DT", "pcaMethods", "statmod")

check_pkgs <- function(x) {
  if(!requireNamespace(x, quietly = TRUE)) {
    stop(paste0("The package ",
                paste0(x),
                " is required to knit this document. 
                Please install it."))
  }
}

lapply(req_pkgs, check_pkgs)
```

# Preface

## Introduction

This vignette outlines the method used to analyse the label-free quantitative 
LC-MS/MS data obtained for _C. albicans_ (strain DAY286 yeast) 
extracellular vesicles (EVs) and whole cell lysates (WCL) by Dawson et 
al. (2019). The tools provided in this package, 
`r Githubpkg("csdaw/candidaev")`, allows the step-by-step reproduction of the 
results presented in the paper. 
The manuscript in its submitted form and the `.Rmarkdown` file used to 
generate the main and supplementary figures can found 
[here](https://github.com/csdaw/candidaev/tree/master/inst/manuscript).

The analysis workflow used in this vignette is derived from the excellent 
**Bioconductor** package `r Biocpkg("DEP")`[^1]. Furthermore, some of the 
plotting functions available in `r Biocpkg("DEP")` have been modified for use 
in this package, `r Githubpkg("csdaw/candidaev")`. 

[^1]: Zhang, X., Smits, A. H., van Tilburg, G. B. A., Ovaa, H., Huber, W., & 
Vermeulen, M. (2018). Proteome-wide identification of ubiquitin interactions 
using UbIA-MS. _Nature Protocols, 13_(3), 530-550. 
[doi:10.1038/nprot.2017.147](https://doi.org/10.1038/nprot.2017.147)

## Experimental design

Extracellular vesicles (EVs) were isolated using differential 
ultracentrifugation from _C. albicans_ liquid cultures 
grown using the lab strain DAY286 (n = 3). 
Additionally, cells from these liquid cultures were separated and lysed to 
generate whole cell lysates (WCL). Each EV biological replicate was paired 
with a WCL replicate as they were derived from the same liquid culture.

Samples analysed in this vignette:

* DAY286 Yeast EV1 (C6.EV)
* DAY286 Yeast EV2 (C7.EV)
* DAY286 Yeast EV3 (C8.EV)
* DAY286 Yeast WCL1 (C6.WCL)
* DAY286 Yeast WCL2 (C7.WCL)
* DAY286 Yeast WCL3 (C8.WCL)

The samples not analysed in this vignette (due to a different EV isolation 
protocol being used) include:

* DAY286 Yeast EV0 (C3.EV)
* DAY286 Yeast WCL0 (C3.WCL)

These samples were electrophoresed and the SDS-PAGE gel slices were 
reduced, alkylated, and digested with trypsin. The resulting peptides were 
extracted with acetonitrile, analysed by nano-LC-ESI-MS/MS, identified using 
MaxQuant, and quantified via the MaxLFQ algorithm in MaxQuant. 
The `proteinGroups.txt` file output was used for subsequent analyses.

## Data availability

The raw data was searched against the _C. albicans_ (SC5314) UniProt 
reference proteome (UP000000559) using MaxQuant version 1.6.0.16.  The 
`proteinGroups.txt` file analysed in this vignette 
can be easily accessed 
[here](https://github.com/csdaw/candidaev/tree/master/data-raw/lcms). 
It is also installed along with the 
`r Githubpkg("csdaw/candidaev")` package and can be loaded in R with 

```
data(yeast)
```

The LC-MS/MS raw data, peak data, and results files have been 
deposited in the ProteomeXchange Consortium database via the PRIDE partner 
repository with the data set identifier PXD014367. 

# Data analysis

## Setup

The packages required for the analyses are loaded.

```{r setup, echo=TRUE, warning=FALSE, message=FALSE}
# load packages
library(candidaev)
library(dplyr)
library(ggplot2)
```

## Exploring the `yeast` data set

As above, the _C. albicans_ DAY286 yeast LC-MS/MS data set is provided as a 
data.frame called `yeast` in this package. 
It has `r nrow(yeast)` rows each of which corresponds to a single protein 
group[^2], and `r ncol(yeast)` columns. 
A key describing most of these columns can be found in the 
[MaxQuant documentation](http://www.coxdocs.org/doku.php?id=maxquant:table:proteingrouptable). 
The `LFQ.intensity` columns hold the protein quantification data that is 
used for the following analyses.

[^2]: Each protein group comprises the group of proteins that could be 
reconstructed from a set of peptides by MaxQuant. In most cases this will be 
a single protein but sometimes there will be two or more proteins that 
could not be distinguished by the set of peptides.

```{r yColumns, echo=TRUE}
colnames(yeast)
```

`yeast_exp` is another data.frame provided in this package. It 
describes the experimental design underlying the `yeast` data set. 
The primary function `yeast_exp` is to provide info for labelling 
and grouping data in some of the plotting functions from 
the **Bioconductor** package `r Biocpkg("DEP")`, which have been 
modified for use in this package. `yeast_exp` supplies sample 
information (biological replicate, strain) for each 
`LFQ.intensity` column in the `yeast` data.frame.

It has the following structure. Notice that each row in the `label` 
column of `yeast_exp` is exactly the same as the names of the 
`LFQ.intensity` columns of `yeast`.

```{r yExpTable}
knitr::kable(yeast_exp, 
             caption = "Experimental design for this vignette.")
```

## Data preparation

First, the `yeast` proteinGroups data.frame is filtered to remove reverse 
proteins, contaminant proteins, and proteins identified by less than 2 unique 
peptides.

```{r yInit, echo=TRUE}
# filter out reverse, contaminant proteins, proteins with <2 unique peptides
# extract LFQ intensity columns and UniProt accessions 
# log2 transform LFQ intensities
yeast_lfq <- yeast %>% 
  filter(Reverse != "+", 
         Potential.contaminant != "+", 
         Unique.peptides >= 2) %>% 
  convert_lfq(., yeast_exp)
```

Then the data.frame is converted into a compact format using the `convert_lfq` 
function which performs the following:

1. Ensures the names of the `LFQ.intensity` columns in `yeast` exactly match 
   those in the `label` column of the experimental design, `yeast_exp`.
   
2. Returns a matrix containing only: the UniProt accession(s) and LFQ 
   intensities of each protein across the 8 samples. 
   
3. Log<sub>2</sub> transforms the LFQ intensities.

--------------------------------------------------------------------------------

Next, the EV0 sample is dropped because the EVs were not isolated 
with the same protocol as the other samples and are hence not equivalent 
biological replicates. The WCL0 sample is dropped to keep the experimental 
design balanced.

```{r yDrop, echo=TRUE}
yeast_lfq2 <- yeast_lfq[, !colnames(yeast_lfq) %in% c("EV_0", "WCL_0")]
```

The resulting matrix `yeast_lfq2` still has `r nrow(yeast_lfq2)` rows but 
only `r ncol(yeast_lfq2)` columns, one for each LC-MS/MS sample. The total
number of proteins quantified (i.e. have a non-`NA` LFQ intensity) in each
sample can be explored using:

```{r, echo=TRUE, eval=FALSE}
# explore quantification across all 6 samples
plot_frequency2(yeast_lfq2)

# explore protein quantification in individual samples
plot_numbers2(yeast_lfq2, yeast_exp)
```

The proteins in this data set 
are not quantified in every biological replicate of every strain. 
If fact in Figure \@ref(fig:yInit2)A shows that most of the proteins are 
only quantified in a single sample. Figure \@ref(fig:yInit2)B shows the 
total number of proteins quantified in each sample.

```{r yInit2, fig.cap=yInit2_cap, fig.wide=TRUE}
freq1 <- plot_frequency2(yeast_lfq2) + 
  scale_y_continuous(limits = c(0,500), 
                     expand = c(0, 0)) + 
  theme(plot.title = element_text(size = 16))

numb1 <- plot_numbers2(yeast_lfq2, yeast_exp) + 
  theme(plot.title = element_text(size = 16),
        axis.text.x = element_text(angle = 0), 
        legend.position = "right", 
        legend.text = element_text(size = 12), 
        legend.box.margin = margin(t = 0, r = 0, b = 0, l = -15, unit = "pt")) +  
  scale_y_continuous(limits = c(0, 1750), 
                     expand = c(0, 0)) + 
  scale_x_discrete(labels = c(rep(2:4, 2))) + 
  labs(title = "Proteins per biological replicate") + 
  xlab("Biological replicate") + 
  guides(fill = guide_legend(nrow = 4, byrow = TRUE))

ggpubr::ggarrange(freq1, numb1,
                  labels = c("A", "B"), 
                  ncol = 2, 
                  nrow = 1, 
                  font.label = list(size = 24, color = "black"), 
                  widths = c(1, 1.4), 
                  vjust = 1)

yInit2_cap <- "Protein quantification distributions. (A) Cumulative 
               number of proteins quantified in 1-6 samples. (B) Total number 
               of proteins quantified in each biological replicate sample."
```

The percentage of non-valid (`NA`) values and other useful statistics for 
each sample is shown using:

```{r, echo=TRUE, eval=FALSE}
# generate summary table from LFQ data
sum_table(yeast_lfq2)
```

Table \@ref(tab:yInit3) shows that biological replicate 2 of both EV and WCL 
has the highest percentage of missing values compared to the other replicates. 

```{r yInit3}
knitr::kable(sum_table(yeast_lfq2), 
             caption = "Descriptive statistics for each LC-MS/MS sample.", 
             col.names = c("Sample", "Mean LFQ intensity", 
                           "Median LFQ intensity", "SD", "CV", 
                           "Min LFQ intensity", "Max LFQ intensity", 
                           "Range", "Number of proteins", "Number of NA", 
                           "Percent NA"), 
             digits = 1)
```

## Protein filtering (replicate quantification)

Prior to normalisation and imputation, proteins with too many missing values 
across the 6 samples are removed. Only those with at least 2/3 valid values 
in 1 of the 2 conditions are retained.

This filtration is performed using `filter_na2`.

```{r yFilt}
# filter for proteins identified in min 2/3 reps of EV or WCL
yeast_filt2 <- filter_na2(yeast_lfq2, 
                          logic = "or", 
                          op = "<=", 
                          pat1 = "EV", val1 = 1, 
                          pat2 = "W", val2 = 1)
```

The initial `r nrow(yeast_lfq2)` proteins has been reduced to 
`r nrow(yeast_filt2)`. Now majority of proteins are quantified in 3/6 or 6/6 
samples. The overall pattern of proteins in each sample remains similar 
though.

```{r yFilt2, fig.cap=yFilt2_cap, fig.wide=TRUE}
freq2 <- plot_frequency2(yeast_filt2) + 
  scale_y_continuous(limits = c(0,400), 
                     expand = c(0, 0)) + 
  theme(plot.title = element_text(size = 16))

numb2 <- plot_numbers2(yeast_filt2, yeast_exp) + 
  theme(plot.title = element_text(size = 16),
        axis.text.x = element_text(angle = 0), 
        legend.position = "right", 
        legend.text = element_text(size = 12), 
        legend.box.margin = margin(t = 0, r = 0, b = 0, l = -15, unit = "pt")) +  
  scale_y_continuous(limits = c(0, 1000), 
                     expand = c(0, 0)) + 
  scale_x_discrete(labels = c(rep(2:4, 2))) + 
  labs(title = "Proteins per biological replicate") + 
  xlab("Biological replicate") + 
  guides(fill = guide_legend(nrow = 4, byrow = TRUE))

ggpubr::ggarrange(freq2, numb2,
                  labels = c("A", "B"), 
                  ncol = 2, 
                  nrow = 1, 
                  font.label = list(size = 24, color = "black"), 
                  widths = c(1, 1.4), 
                  vjust = 1)

yFilt2_cap <- "Protein quantification distributions after filtration. 
               (A) Cumulative number of proteins quantified in 1-6 samples. 
               (B) Total number of proteins quantified in each biological 
               replicate sample."
```

A multidimensional scaling plot can be drawn via:

```{r, echo=TRUE, eval=FALSE}
# need labels for the MDS plot
mds_labels <- c(rep("EV", 3), 
                rep("WCL", 3))

# generate MDS plot
plot_mds(yeast_filt2, mat_labels = mds_labels, 
         shape_size = 4)
```

It shows that the EV samples generally cluster closer together than 
the WCL samples do to each other, indicating they have a smaller 
log<sub>2</sub> fold change between them.

```{r bFilt3, fig.cap=bFilt3_cap, fig.small=TRUE}
mds1_labels <- c(rep("EV", 3), 
                 rep("WCL", 3))

plot_mds(yeast_filt2, mat_labels = mds1_labels, 
         shape_size = 4) + 
  scale_y_continuous(limits = c(-1.5, 2), 
                     expand = c(0, 0)) + 
  scale_x_continuous(limits = c(-1, 3), 
                     expand = c(0, 0)) + 
  theme(legend.position = "bottom", 
        legend.title = element_blank()) + 
  scale_fill_manual(values = c("orange", "red")) + 
  xlab("Dimension 1") + 
  ylab("Dimension 2")

bFilt3_cap <- "Multidimensional scaling (MDS) plot. Shows each LC-MS/MS
               biological replicate sample with the first 2 dimensions 
               plotted. Pairwise distance between samples approximately 
               indicates the log2 fold change between samples."
```

## LFQ intensity normalisation

Next, the LFQ intensity data is normalised using `r Biocpkg("limma")`.

```{r yNorm, echo=TRUE}
# normalise LFQ intensities
yeast_norm2 <- limma::normalizeCyclicLoess(yeast_filt2)
```

The effect of the normalisation on the distribution of LFQ intensities for 
each sample is visualised using: 

```{r, echo=TRUE, eval=FALSE}
# explore effect of normalisation on intensity distributions
plot_normalization2(yeast_exp,
                    facet_labels = c(yeast_filt2 = "Before normalisation", 
                                     yeast_norm2 = "After normalisation"), 
                    yeast_filt2, 
                    yeast_norm2)
```

```{r, yNorm2, fig.cap=bNorm2_cap, fig.height=7}
plot_normalization2(yeast_exp, 
                    facet_labels = c(yeast_filt2 = "Before normalisation", 
                                     yeast_norm2 = "After normalisation"), 
                    yeast_filt2, yeast_norm2) + 
  scale_x_discrete(labels = c(rep(2:4, 2))) + 
  xlab("Biological replicate") + 
  theme(plot.margin = margin(t = 0, l = 1.5, r = 1.5, b = 0, unit = "inches"))
  

bNorm2_cap <- "Distribution of protein LFQ intensities for each sample. Shown
               before and after normalisation."
```

The MDS plot now shows that the EV samples and WCL samples cluster separately 
from each other. The first dimension (which is the 
leading factor causing the LFQ intensities differences between samples) 
separates the samples according to the primary condition of interest; EV 
versus WCL.

```{r yNorm3, fig.cap=yNorm3_cap, fig.small=TRUE}
mds2_labels <- c(rep("EV", 3), 
                rep("WCL", 3))

plot_mds(yeast_norm2, mat_labels = mds2_labels, 
         shape_size = 4) + 
  scale_y_continuous(limits = c(-0.5, 0.5), 
                     expand = c(0, 0)) + 
  scale_x_continuous(limits = c(-1.25, 1.25), 
                     expand = c(0, 0)) + 
  theme(legend.position = "bottom", 
        legend.title = element_blank()) + 
  scale_fill_manual(values = c("orange", "red")) + 
  xlab("Dimension 1") + 
  ylab("Dimension 2")

yNorm3_cap <- "Multidimensional scaling (MDS) plot. Shows each LC-MS/MS
               biological replicate sample after normalisation, with the 
               first 2 dimensions plotted. Pairwise distance between samples 
               approximately indicates the log2 fold change between samples."
```

## Yeast EV versus WCL protein comparison

Lists of proteins quantified in at least 1/3 biological replicates of each 
condition are generated using:

```{r yVennData, echo=TRUE}
# filter for proteins quantified in min 1/3 reps of EV
y_ev <- filter_na(yeast_norm2, 
                  op = "<=", 
                  pat = "EV", 
                  val = 2)

# filter for proteins quantified in min 1/3 reps of WCL
y_wcl <- filter_na(yeast_norm2, 
                   op = "<=", 
                   pat = "W", 
                   val = 2)
```

These are then compared via a Venn diagram. As per section 2.4, every protein 
that appears in the Venn diagram has been quantified in at least 2/3 
replicates of EV or WCL.

```{r, echo=TRUE, eval=FALSE}
# define a comparison list for a Venn diagram
venn_comp <- list("EV" = rownames(y_ev), 
                  "WCL" = rownames(y_wcl))

# plot the Venn diagram
venn <- plot_venn(venn_comp, 
                  type = "plot")

grid::grid.draw(venn)
```

Of `r nrow(yeast_norm2)` proteins, 494 (51.7%) are quantified by at least 1/3 
biological replicates of both EV and WCL **and** at least 2/3 replicates of 
either EV or WCL. 

```{r yVenn, fig.cap=venn_cap, fig.small=TRUE}
# define comparison list for Venn diagram
venn1_comp <- list("EV" = rownames(y_ev), 
                   "WCL" = rownames(y_wcl))

# define ggplot2 default colour palette function
gg_colour <- function(n) {
  hues <- seq(15, 375, length = n + 1)
  grDevices::hcl(h = hues, l = 65, c = 100)[1:n]
}

# plot Venn diagram
venn1 <- plot_venn(venn1_comp, 
                   use_uniprot = FALSE, 
                   type = "plot", 
                   fontfamily = "sans", 
                   cat.fontfamily = "sans", 
                   cex = 1.9, 
                   cat.cex = 1.2, 
                   fill = gg_colour(2), 
                   alpha = c(rep(0.6, 2)))

# draw Venn diagram
grid::grid.draw(venn1)

venn_cap <- "Venn diagram of proteins quantified in each condition. 
             Proteins in this plot are quantified in at least 2/3 replicates 
             of either EV or WCL."
```

Next, the most abundant proteins that are only detected in DAY286 yeast 
EVs and not in the WCL can be identified.

```{r yTop10Data}
# get top 10 most abundant protein exclusive to yeast EVs
# in descending order of abundance
yeast_top10 <- y_ev %>% 
  tibble::as_tibble(rownames = "id") %>% 
  filter_na(., op = "==", pat = "W", val = 3) %>% 
  select(-contains("W")) %>% 
  mutate(mean = rowMeans(select(., contains("EV")), na.rm = TRUE)) %>% 
  top_n(n = 10, wt = mean) %>% 
  arrange(desc(mean)) %>% 
  match_id_multi(., "id", uniprot, "UP_accession", 
                 new = c("CGD_gene_name", 
                         "CGD_description")) %>% 
  select(CGD_gene_name, CGD_description, everything(), -id)
```

```{r yTop10}
knitr::kable(yeast_top10, 
             caption = "Top 10 most abundant DAY286 yeast EV proteins. They 
                        are not detected in DAY286 yeast WCL.", 
             col.names = c("Name", "Function", "EV2 LFQ", 
                           "EV3 LFQ", "EV4 LFQ", "Mean LFQ"), 
             digits = 1)
```

## Imputation of missing values

Missing values in the data will be imputed prior to the 
differential abundance analysis. Before imputation the percentage of missing 
values in each sample is explored again with `sum_table(yeast_norm2)`.

```{r ySum}
# explore percentage missing values in each sample
knitr::kable(sum_table(yeast_norm2), 
             caption = "Descriptive statistics for each LC-MS/MS sample.", 
             col.names = c("Sample", "Mean LFQ intensity", 
                           "Median LFQ intensity", "SD", "CV", 
                           "Min LFQ intensity", "Max LFQ intensity", 
                           "Range", "Number of proteins", "Number of NA", 
                           "Percent NA"), 
             digits = 1)
```

The percentage of missing values ranges from 21.9% up to 49.4% which is 
quite high, typical of label-free quantitative proteomics data[^3]. The 
pattern of missing values across the data is visualised using:

[^3]: Lazar, C., Gatto, L., Ferro, M., Bruley, C., Burger, T. (2016). 
Accounting for the Multiple Natures of Missing Values in Label-Free 
Quantitative Proteomics Data Sets to Compare Imputation Strategies. 
_Journal of Proteome Research, 15_(4), 1116-1125. 
[doi:10.1021/acs.jproteome.5b00981](https://doi.org/10.1021/acs.jproteome.5b00981)

```{r, echo=TRUE, eval=FALSE}
# visualise missing values using a heatmap
plot_missval2(yeast_norm2)
```

```{r yMissVal, fig.cap=missval_cap, fig.height=8}
plot_missval2(yeast_norm2)

missval_cap <- "Pattern of missing LFQ intensity values."
```

There are large chunks of data missing in either all of 
the WCL samples or all of the EV samples. To minimise the percentage of 
missing values in each sample, these proteins are removed prior to 
imputation and hence are not imputed.

```{r yExcl, echo=TRUE}
# should not impute with so many missing values
# select proteins with 3/3 missing values in EV or WCL
# these will not be imputed
yeast_excl <- filter_na2(yeast_norm2, 
                         logic = "or", 
                         op = "==", 
                         pat1 = "EV", val1 = 3, 
                         pat2 = "W", val2 = 3)
```

Only proteins quantified in at least 1/3 replicates of both EV and WCL will be 
retained for imputation.

```{r yBoth, echo=TRUE}
# filter for proteins with at least 1 valid value in EV and WCL
yeast_both <- filter_na2(yeast_norm2, 
                         logic = "and", 
                         op = "<=", 
                         pat1 = "EV", val1 = 2, 
                         pat2 = "W", val2 = 2)
```

`r nrow(yeast_excl)` proteins which have 3/3 missing values in EV or 
WCL are removed (`yeast_excl`). 
Conversely, `r nrow(yeast_both)` proteins are kept to have missing values 
imputed (`yeast_both`). The percentage of missing values across each sample 
in `yeast_both` now ranges from 2.2% to 31.4%. 

```{r ySum2}
knitr::kable(sum_table(yeast_both), 
             caption = "Descriptive statistics for each LC-MS/MS sample.",
             col.names = c("Sample", "Mean LFQ intensity", 
                           "Median LFQ intensity", "SD", "CV", 
                           "Min LFQ intensity", "Max LFQ intensity", 
                           "Range", "Number of proteins", "Number of NA", 
                           "Percent NA"), 
             digits = 1)
```

The pattern of missing values in `yeast_both` can be visualised.

```{r yMissVal2, fig.cap=missval2_cap , fig.height=8}
plot_missval2(yeast_both)

missval2_cap <- "Pattern of missing LFQ intensity values."
```

The LFQ intensity distributions and cumulative fraction of proteins with and 
without missing values can be plotted with: 

```{r, echo=TRUE, eval=FALSE}
# detect intensity differences for proteins with missing values
plot_detect2(yeast_both)
```

```{r bSum3, fig.cap=detect_cap, fig.wide=TRUE}
plot_detect2(yeast_both)

detect_cap <- "LFQ intensities of proteins with and without missing values."
```

Missing values are imputed using the `bpca` method in the `pca` function from 
the **Bioconductor** package `r Biocpkg("pcaMethods")`, which is designed for 
missing value imputation.

```{r, echo=TRUE, eval=FALSE}
# impute missing values with bpca method
yeast_imp <- pcaMethods::pca(yeast_both, method = "bpca")
yeast_imp <- pcaMethods::completeObs(yeast_imp)
```

```{r, echo=TRUE, eval=FALSE}
plot_imputation2(yeast_exp, 
                 facet_labels = c(yeast_both = "Before imputation", 
                                  yeast_imp = "After imputation"), 
                 yeast_both, 
                 yeast_imp)
```

```{r yImpute, fig.cap=imp_cap, fig.wide=TRUE}
# impute missing values with bpca method
yeast_imp <- pcaMethods::pca(yeast_both, method = "bpca")
yeast_imp <- pcaMethods::completeObs(yeast_imp)

plot_imputation2(yeast_exp, 
                 facet_labels = c(yeast_both = "Before imputation", 
                                  yeast_imp = "After imputation"), 
                 yeast_both, yeast_imp)

imp_cap <- "LFQ intensity distribution before and after imputation."
```

## Differential abundance analysis with _limma_ 

Prior to differential abundance analysis, the imputed and non-imputed proteins 
are recombined in the same matrix.

```{r yDE, echo=TRUE}
# combine matrices
yeast_de <- rbind(yeast_excl, yeast_imp)
```

A design matrix is constructed for this proteomics experiment. This is 
required for the linear model fit function (`lmFit`) from 
`r Biocpkg("limma")` which is called inside the `limma_eBayes` function in 
this package. See the _limma_ User Guide section 9.2
[here](https://bioconductor.org/packages/devel/bioc/vignettes/limma/inst/doc/usersguide.pdf) 
for more information about constructing design matrices.

```{r yDesign, echo=TRUE}
# see limma user guide section 9.2 for more info about DE
# create design matrix
y_samp <- data.frame(T = (rep(c("EV", "WCL"), each = 3)))

y_design <- stats::model.matrix(~ 0 + T, data = y_samp)
colnames(y_design) <- c("EV", "WCL")
```

```{r yDE2}
knitr::kable(y_design, 
             caption = "Design matrix for comparing EV and WCL samples.")
```

The comparisons between samples that one is interested in are defined in a 
character vector. These must be two column names from the design matrix 
separated by ` - `.

```{r yDE3, echo=TRUE}
# specify comparisons of interest
y_contrasts <- c("EV - WCL")
```

Finally, the differential abundance analysis is performed using the 
`limma_eBayes` function from this package which calls `eBayes` from 
`r Biocpkg("limma")`. 

To perform the analysis, the following objects are provided:

* a numeric matrix of log<sub>2</sub> LFQ intensity data (`yeast_de`).
* a numeric design matrix (`y_design`).
* a character vector with the contrasts of interest (`y_contrasts`).

The results from `limma_eBayes` are extracted using `get_results` 
which allows the adjusted p-value cutoff and 
log<sub>2</sub> fold change cutoff for significance to be specified. To get the 
overall differences for all the samples included in the design matrix, 
`type = "overall"` is used. However in this case there is only one contrast of 
interest in which case `type = "individual"` should be used to obtain the 
log<sub>2</sub> fold changes and adjusted p-values 
for this comparison. The result table is accessed by subsetting the list of 
data.frames output by `get_results`.

```{r yDE4, echo=TRUE, warning=FALSE}
# make all pair-wise comparisons between EV and WCL
# and perform limma::eBayes()
y_efit <- limma_eBayes(yeast_de, y_design, y_contrasts)

# extract DE results
yeast_res <- get_results(efit = y_efit, 
                         mat = yeast_de, 
                         p_val = 0.01, 
                         lfc = 0, 
                         type = "individual")[[1]]
```

# Results

## DAY286 yeast EV versus WCL

Differential abundance comparison of proteins in DAY286 yeast EVs versus WCL.

```{r yResData, include=FALSE}
yeast_res_short <- yeast_res %>% 
  select(CGD_gene_name, Protein_name, logFC, 
         contains("_mean"), adj.P.Val, group) %>% 
  mutate_at(.vars = vars(logFC, contains("_mean")), 
            .funs = list(~round(., digits = 2))) %>% 
  mutate(adj.P.Val = signif(adj.P.Val, digits = 3), 
         group = as.factor(group))

# caption for volcano plot below
vol_cap <- "Volcano plot for proteins quantified in EV and WCL."
```

```{r yRes}
js1 <- DT::JS("function(data, type, row, meta) {",
              "return type === 'display' && data.length > 40 ?",
              "'<span title=\"' + data + '\">' + data.substr(0, 40) + 
              '...</span>' : data;",
              "}")

js2 <- DT::JS("function(data, type, row, meta) {",
              "return type === 'display' && data.length > 12 ?",
              "'<span title=\"' + data + '\">' + data.substr(0, 12) + 
              '...</span>' : data;",
              "}")

DT::datatable(yeast_res_short, 
              rownames = FALSE, 
              colnames = c("Gene" = "CGD_gene_name", 
                           "Protein" = "Protein_name", 
                           "p.adj" = "adj.P.Val"), 
              filter = list(position = "bottom", clear = FALSE, plain = TRUE), 
              class = "display compact", 
              options = list(
                autowidth = TRUE, 
                columnDefs = list(list(width = "40%", targets = 1), 
                                  list(width = "10%", targets = c(0, 2:6)), 
                                  list(targets = 1, 
                                       render = js1), 
                                  list(targets = 0, 
                                       render = js2)
                )
              )
)
```

## Volcano plot

The result table is used to construct a volcano plots for DAY286 yeast. 
The `plot_volcano` function requires a data.frame input (`yeast_res`) 
and specification of the column names that contain the data to be plotted 
(`adj.P.Val` and `logFC`). The `group` column in `yeast_res` is used to 
separate the data into three groups that have different colours.

```{r yVolcano, fig.cap=vol_cap, echo=TRUE}
plot_volcano(yeast_res, 
             p_val = adj.P.Val, 
             log2fc = logFC, 
             group = group, 
             use_labels = TRUE, 
             label_fun = ifelse(logFC >= 6, CGD_gene_name, ""), 
             point_size = 2, 
             label_size = 4, 
             legend_params = c("right", 14, 14), 
             axis_params = c("black", 16, 14), 
             x_lim = c(-10, 13), 
             y_lim = c(0, 7)) + 
  scale_colour_manual(values = c("#D55E00",
                                 "#0072B2",
                                 "#999999"), 
                      labels = c("EV enriched", 
                                 "WCL enriched", 
                                 "Not significant")) + 
  theme(legend.title = element_blank())
```

## Log<sub>2</sub> LFQ intensity heatmap

The `yeast_res` data.frame is used to construct heatmaps based on the 
proteins that are significantly different across all both sample groups; 
EV and WCL.

There are two type of heatmaps: one for plotting log<sub>2</sub> LFQ intensity 
data and one for plotting log<sub>2</sub> fold change data. To plot an LFQ 
intensity heatmap, a numeric matrix must be used as the input.

In this case the `yeast_res` data.frame is subset to remove all columns 
except a column with identifiers (e.g. UniProt accessions), 
log2 LFQ intensity data columns, and a logical significance column. 
Then the data.frame is 
filtered for significant proteins, the UniProt accession column is converted to 
rownames, the significance column is dropped, and the data.frame is converted 
to a matrix.

```{r lfqHeatmapData, echo=TRUE}
# get data for log2 LFQ intensity heatmap
yeast_htmp <- yeast_res %>% 
  select(UP_accession, 
         matches("EV_[0-9]"), 
         matches("WCL_[0-9]"), 
         significant) %>% 
  filter(significant == TRUE) %>% 
  tibble::column_to_rownames(var = "UP_accession") %>% 
  select(-significant) %>% 
  as.matrix()
```

The heatmap is plotted with: 

```{r, echo=TRUE, eval=FALSE}
htmp_params <- list(title = expression("centered"~log[2]~"LFQ intensity"), 
                    color_bar = "continuous", 
                    direction = "vertical", 
                    legend_height = unit(10, "cm"), 
                    title_position = "lefttop-rot", 
                    title_gp = grid::gpar(colour = "black", fontsize = 16), 
                    labels_gp = grid::gpar(colour = "black", fontsize = 16), 
                    at = seq(-5, 5, 1))

plot_heatmap(yeast_htmp, 
             plot = TRUE, 
             data_type = "log2intensity", 
             clust_fun = "gower", 
             split_type = "cutree", 
             k = 5, 
             legend_pos = "right",
             cluster_split = FALSE, 
             split_order = c(4, 3, 1, 2, 5), 
             colour_lims = c(-4,4), 
             row_title_fontsize = 16, 
             col_name_fontsize = 14, 
             heatmap_legend_param = htmp_params)
```

```{r lfqHeatmap, fig.cap=htmp_cap, fig.height=8}
htmp_params <- list(title = expression("centered"~log[2]~"LFQ intensity"), 
                    color_bar = "continuous", 
                    direction = "vertical", 
                    legend_height = unit(10, "cm"), 
                    title_position = "lefttop-rot", 
                    title_gp = grid::gpar(colour = "black", fontsize = 16), 
                    labels_gp = grid::gpar(colour = "black", fontsize = 16), 
                    at = seq(-5, 5, 1))

plot_heatmap(yeast_htmp, 
             plot = TRUE, 
             data_type = "log2intensity", 
             clust_fun = "gower", 
             split_type = "cutree", 
             k = 5, 
             legend_pos = "right",
             cluster_split = FALSE, 
             split_order = c(4, 3, 1, 2, 5), 
             colour_lims = c(-4,4), 
             row_title_fontsize = 16, 
             col_name_fontsize = 14, 
             heatmap_legend_param = htmp_params)

htmp_cap <- "Protein log<sub>2</sub> LFQ intensity heatmap. Data has been 
             centred around 0."
```

# Session information {-}

The system on which this document was compiled was running 
pandoc v`r rmarkdown::pandoc_version()`. Here is output of 
`sessionInfo()`.

```{r sessionInfo}
sessionInfo()
```
