---
title: "Vignette title"
author: 
- name: Charlotte Dawson
  affiliation: Department of Biochemistry and Genetics, 
               La Trobe Institute of Molecular Science, 
               La Trobe University, Australia
- name: Mark Bleackley
  affiliation: Department of Biochemistry and Genetics, 
               La Trobe Institute of Molecular Science, 
               La Trobe University, Australia
- name: Donovan Garcia Ceron
  affiliation: Department of Biochemistry and Genetics, 
               La Trobe Institute of Molecular Science, 
               La Trobe University, Australia
- name: Marilyn Anderson
  affiliation: Department of Biochemistry and Genetics, 
               La Trobe Institute of Molecular Science, 
               La Trobe University, Australia
date: '`r format(Sys.Date(), "%d %B %Y")`'
output:
  BiocStyle::html_document:
    toc: true
    df_print: kable
package: candidaev
abstract: | 
  This is the abstract.
vignette: > 
  %\VignetteIndexEntry{Vignette title}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r knitrOpts, include=FALSE}
# knit options
knitr::opts_knit$set(progress = TRUE, 
                     verbose = TRUE)

# chunk options
knitr::opts_chunk$set(collapse = FALSE,
                      warning = TRUE,
                      message = TRUE,
                      echo = FALSE,
                      error = TRUE, 
                      comment = "#>")

# check for installation of packages requires to knit this vignette
req_pkgs <- c("BiocStyle", "ggpubr", "stringi", "DT")

check_pkgs <- function(x) {
  if(!requireNamespace(x, quietly = TRUE)) {
    stop(paste0("The package ",
                paste0(x),
                " is required to knit this document. 
                Please install it."))
  }
}

lapply(req_pkgs, check_pkgs)
```

# Preface

## Introduction

This vignette outlines the method used to analyse the label-free quantitative 
LC-MS/MS data obtained for _C. albicans_ (strain DAY286 yeast) 
extracellular vesicles (EVs) and whole cell lysates (WCL) by Dawson et 
al. (2019). The tools provided in this package, 
`r Githubpkg("csdaw/candidaev")`, allows the step-by-step reproduction of the 
results presented in the paper. 
The manuscript in its submitted form and the `.Rmarkdown` file used to 
generate it can found 
[here](https://github.com/csdaw/candidaev/tree/master/manuscript).

The analysis workflow used in this vignette is derived from the excellent 
**Bioconductor** package `r Biocpkg("DEP")`[^1]. Furthermore, some of the 
plotting functions available in `r Biocpkg("DEP")` have been modified for use 
in this package, `r Githubpkg("csdaw/candidaev")`. 

[^1]: Zhang, X., Smits, A. H., van Tilburg, G. B. A., Ovaa, H., Huber, W., & 
Vermeulen, M. (2018). Proteome-wide identification of ubiquitin interactions 
using UbIA-MS. _Nature Protocols, 13_(3), 530-550. 
[doi:10.1038/nprot.2017.147](https://doi.org/10.1038/nprot.2017.147)

## Experimental design

Extracellular vesicles (EVs) were isolated using differential 
ultracentrifugation from _C. albicans_ liquid cultures 
grown using the lab strain DAY286 (n = 3). 
Additionally, cells from these liquid cultures were separated and lysed to 
generate whole cell lysates (WCL). Each EV biological replicate was paired 
with a WCL replicate as they were derived from the same liquid culture.

Samples analysed in this vignette:

* DAY286 Yeast EV2 (C6.EV)
* DAY286 Yeast EV3 (C7.EV)
* DAY286 Yeast EV4 (C8.EV)
* DAY286 Yeast WCL2 (C6.WCL)
* DAY286 Yeast WCL3 (C7.WCL)
* DAY286 Yeast WCL4 (C8.WCL)

The samples not analysed in this vignette (due to a different EV isolation 
protocol being used) include:

* DAY286 Yeast EV1 (C3.EV)
* DAY286 Yeast WCL1 (C3.WCL)

These samples were electrophoresed and the SDS-PAGE gel slices were 
reduced, alkylated, and digested with trypsin. The resulting peptides were 
extracted with acetonitrile, analysed by nano-LC-ESI-MS/MS, identified using 
MaxQuant, and quantified via the MaxLFQ algorithm in MaxQuant. 
The `proteinGroups.txt` file output was used for subsequent analyses.

## Data availability

The raw data was searched against the _C. albicans_ (SC5314) UniProt 
reference proteome (UP000000559) using MaxQuant version 1.6.0.16.  The 
`proteinGroups.txt` file analysed in this vignette 
can be easily accessed 
[here](https://github.com/csdaw/candidaev/tree/master/data-raw/lcms). 
It is also installed along with the 
`r Githubpkg("csdaw/candidaev")` package and can be loaded in R with 

```
data(yeast)
```

The LC-MS/MS raw data, peak data, and results files have been 
deposited in the ProteomeXchange Consortium database via the PRIDE partner 
repository with the data set identifier y and DOI x. 

# Data analysis

## Setup

The packages required for the analyses are loaded and a specific 
seed is set for the steps requiring random number generation (imputation 
and clustering).

```{r setup, echo=TRUE, warning=FALSE, message=FALSE}
# load packages
library(candidaev)
library(dplyr)
library(ggplot2)

# set seed for imputation, kmeans, MDS plot
set.seed(1)
```

## Exploring the `yeast` data set

As above, the _C. albicans_ DAY286 yeast LC-MS/MS data set is provided as a 
data.frame called `yeast` in this package. 
It has `r nrow(yeast)` rows each of which corresponds to a single protein 
group[^2], and `r ncol(yeast)` columns. 
A key describing most of these columns can be found in the 
[MaxQuant documentation](http://www.coxdocs.org/doku.php?id=maxquant:table:proteingrouptable). 
The `LFQ.intensity` columns hold the protein quantification data that is 
used for the following analyses.

[^2]: Each protein group comprises the group of proteins that could be 
reconstructed from a set of peptides by MaxQuant. In most cases this will be 
a single protein but sometimes there will be two or more proteins that 
could not be distinguished by the set of peptides.

```{r yColumns, echo=TRUE}
colnames(yeast)
```

`yeast_exp` is another data.frame provided in this package. It 
describes the experimental design underlying the `yeast` data set. 
The primary function `yeast_exp` is to provide info for labelling 
and grouping data in some of the plotting functions from 
the **Bioconductor** package `r Biocpkg("DEP")`, which have been 
modified for use in this package. `yeast_exp` supplies sample 
information (biological replicate, strain) for each 
`LFQ.intensity` column in the `yeast` data.frame.

It has the following structure. Notice that each row in the `label` 
column of `yeast_exp` is exactly the same as the names of the 
`LFQ.intensity` columns of `yeast`.

```{r yExpTable}
knitr::kable(yeast_exp, 
             caption = "Experimental design for this vignette.")
```

## Data preparation

First, the `yeast` proteinGroups data.frame is filtered to remove reverse 
proteins, contaminant proteins, and proteins identified by less than 2 unique 
peptides.

```{r yInit, echo=TRUE}
# filter out reverse, contaminant proteins, proteins with <2 unique peptides
# extract LFQ intensity columns and UniProt accessions 
# log2 transform LFQ intensities
yeast_lfq <- yeast %>% 
  filter(Reverse != "+", 
         Potential.contaminant != "+", 
         Unique.peptides >= 2) %>% 
  convert_lfq(., yeast_exp)
```

Then the data.frame is converted into a compact format using the `convert_lfq` 
function which performs the following:

1. Ensures the names of the `LFQ.intensity` columns in `yeast` exactly match 
   those in the `label` column of the experimental design, `yeast_exp`.
   
2. Returns a matrix containing only: the UniProt accession(s) and LFQ 
   intensities of each protein across the 8 samples. 
   
3. Log<sub>2</sub> transforms the LFQ intensities.

--------------------------------------------------------------------------------

Next, the EV1 sample is dropped because the EVs were not isolated 
with the same protocol as the other samples and are hence not equivalent 
biological replicates. The WCL1 sample is dropped to keep the experimental 
design balanced.

```{r yDrop, echo=TRUE}
yeast_lfq2 <- yeast_lfq[, !colnames(yeast_lfq) %in% c("EV_1", "WCL_1")]
```

The resulting matrix `yeast_lfq2` still has `r nrow(yeast_lfq2)` rows but 
only `r ncol(yeast_lfq2)` columns, one for each LC-MS/MS sample. The total
number of proteins quantified (i.e. have a non-`NA` LFQ intensity) in each
sample can be explored using:

```{r, echo=TRUE, eval=FALSE}
# explore quantification across all 6 samples
plot_frequency2(yeast_lfq2)

# explore protein quantification in individual samples
plot_numbers2(yeast_lfq2, yeast_exp)
```

The proteins in this data set 
are not quantified in every biological replicate of every strain. 
If fact in Figure \@ref(fig:yInit2)A shows that most of the proteins are 
only quantified in a single sample. Figure \@ref(fig:yInit2)B shows the 
total number of proteins quantified in each sample.

```{r yInit2, fig.cap=yInit2_cap, fig.wide=TRUE}
freq1 <- plot_frequency2(yeast_lfq2) + 
  scale_y_continuous(limits = c(0,500), 
                     expand = expand_scale(mult = c(0, 0))) + 
  theme(plot.title = element_text(size = 16))

numb1 <- plot_numbers2(yeast_lfq2, yeast_exp) + 
  theme(plot.title = element_text(size = 16),
        axis.text.x = element_text(angle = 0), 
        legend.position = "right", 
        legend.text = element_text(size = 12), 
        legend.box.margin = margin(t = 0, r = 0, b = 0, l = -15, unit = "pt")) +  
  scale_y_continuous(limits = c(0, 1750), 
                     expand = expand_scale(mult = c(0, 0))) + 
  scale_x_discrete(labels = c(rep(2:4, 2))) + 
  labs(title = "Proteins per biological replicate") + 
  xlab("Biological replicate") + 
  guides(fill = guide_legend(nrow = 4, byrow = TRUE))

ggpubr::ggarrange(freq1, numb1,
                  labels = c("A", "B"), 
                  ncol = 2, 
                  nrow = 1, 
                  font.label = list(size = 24, color = "black"), 
                  widths = c(1, 1.4), 
                  vjust = 1)

bInit2_cap <- "Protein quantification distributions. (A) Cumulative 
               number of proteins quantified in 1-6 samples. (B) Total number 
               of proteins quantified in each biological replicate sample."
```

The percentage of non-valid (`NA`) values and other useful statistics for 
each sample is shown using:

```{r, echo=TRUE, eval=FALSE}
# generate summary table from LFQ data
sum_table(yeast_lfq2)
```

Table \@ref(tab:yInit3) shows that biological replicate 2 of both EV and WCL 
has the highest percentage of missing values compared to the other replicates. 

```{r yInit3}
knitr::kable(sum_table(yeast_lfq2), 
             caption = "Descriptive statistics for each LC-MS/MS sample.", 
             col.names = c("Sample", "Mean LFQ intensity", 
                           "Median LFQ intensity", "SD", "CV", 
                           "Min LFQ intensity", "Max LFQ intensity", 
                           "Range", "Number of proteins", "Number of NA", 
                           "Percent NA"), 
             digits = 1)
```

## Protein filtering (replicate quantification)

Prior to normalisation and imputation, proteins with too many missing values 
across the 6 samples are removed. Only those with at least 2/3 valid values 
in 1 of the 2 conditions are retained.

This filtration is performed using `filter_na2`.

```{r yFilt}
# filter for proteins identified in min 2/3 reps of EV or WCL
yeast_filt2 <- filter_na2(yeast_lfq2, 
                          logic = "or", 
                          op = "<=", 
                          pat1 = "EV", val1 = 1, 
                          pat2 = "W", val2 = 1)
```

The initial `r nrow(yeast_lfq2)` proteins has been reduced to 
`r nrow(yeast_filt2)`. Now the majority of proteins are quantified in all 6 
samples. The overall pattern of proteins in each sample remains similar 
though.

```{r yFilt2, fig.cap=yFilt2_cap, fig.wide=TRUE}
freq2 <- plot_frequency2(yeast_filt2) + 
  scale_y_continuous(limits = c(0,300), 
                     expand = expand_scale(mult = c(0, 0))) + 
  theme(plot.title = element_text(size = 16))

numb2 <- plot_numbers2(yeast_filt2, yeast_exp) + 
  theme(plot.title = element_text(size = 16),
        axis.text.x = element_text(angle = 0), 
        legend.position = "right", 
        legend.text = element_text(size = 12), 
        legend.box.margin = margin(t = 0, r = 0, b = 0, l = -15, unit = "pt")) +  
  scale_y_continuous(limits = c(0, 1000), 
                     expand = expand_scale(mult = c(0, 0))) + 
  scale_x_discrete(labels = c(rep(2:4, 2))) + 
  labs(title = "Proteins per biological replicate") + 
  xlab("Biological replicate") + 
  guides(fill = guide_legend(nrow = 4, byrow = TRUE))

ggpubr::ggarrange(freq2, numb2,
                  labels = c("A", "B"), 
                  ncol = 2, 
                  nrow = 1, 
                  font.label = list(size = 24, color = "black"), 
                  widths = c(1, 1.4), 
                  vjust = 1)

yFilt2_cap <- "Protein quantification distributions after filtration. 
               (A) Cumulative number of proteins quantified in 1-6 samples. 
               (B) Total number of proteins quantified in each biological 
               replicate sample."
```

A multidimensional scaling plot can be drawn via:

```{r, echo=TRUE, eval=FALSE}
# need labels for the MDS plot
mds_labels <- c(rep("EV", 3), 
                rep("WCL", 3))

# generate MDS plot
plot_mds(yeast_filt2, mat_labels = mds_labels, 
         shape_size = 4)
```

It shows that the EV samples generally cluster closer together than 
the WCL samples do to each other, indicating they have a smaller 
log<sub>2</sub> fold change between them.

```{r bFilt3, fig.cap=bFilt3_cap, fig.small=TRUE}
mds1_labels <- c(rep("EV", 3), 
                 rep("WCL", 3))

plot_mds(yeast_filt2, mat_labels = mds1_labels, 
         shape_size = 4) + 
  scale_y_continuous(limits = c(-1.5, 2), 
                     expand = expand_scale(mult = c(0, 0))) + 
  scale_x_continuous(limits = c(-1, 3), 
                     expand = expand_scale(mult = c(0, 0))) + 
  theme(legend.position = "bottom", 
        legend.title = element_blank()) + 
  scale_fill_manual(values = c("orange", "red")) + 
  xlab("Dimension 1") + 
  ylab("Dimension 2")

bFilt3_cap <- "Multidimensional scaling (MDS) plot. Shows each LC-MS/MS
               biological replicate sample with the first 2 dimensions 
               plotted. Pairwise distance between samples approximately 
               indicates the log2 fold change between samples."
```

## LFQ intensity normalisation

Next, the LFQ intensity data is normalised using `r Biocpkg("limma")`.

```{r yNorm, echo=TRUE}
# normalise LFQ intensities
yeast_norm2 <- limma::normalizeCyclicLoess(yeast_filt2)
```

The effect of the normalisation on the distribution of LFQ intensities for 
each sample is visualised using: 

```{r, echo=TRUE, eval=FALSE}
# explore effect of normalisation on intensity distributions
plot_normalization2(yeast_exp,
                    facet_labels = c(yeast_filt2 = "Before normalisation", 
                                     yeast_norm2 = "After normalisation"), 
                    yeast_filt2, 
                    yeast_norm2)
```

```{r, yNorm2, fig.cap=bNorm2_cap, fig.height=7}
plot_normalization2(yeast_exp, 
                    facet_labels = c(yeast_filt2 = "Before normalisation", 
                                     yeast_norm2 = "After normalisation"), 
                    yeast_filt2, yeast_norm2) + 
  scale_x_discrete(labels = c(rep(2:4, 2))) + 
  xlab("Biological replicate") + 
  theme(plot.margin = margin(t = 0, l = 1.5, r = 1.5, b = 0, unit = "inches"))
  

bNorm2_cap <- "Distribution of protein LFQ intensities for each sample. Shown
               before and after normalisation."
```

The MDS plot now shows that the EV samples and WCL samples cluster separately 
from each other. The first dimension (which is the 
leading factor causing the LFQ intensities differences between samples) 
separates the samples according to the primary condition of interest; EV 
versus WCL.

```{r yNorm3, fig.cap=yNorm3_cap, fig.small=TRUE}
mds2_labels <- c(rep("EV", 3), 
                rep("WCL", 3))

plot_mds(yeast_norm2, mat_labels = mds2_labels, 
         shape_size = 4) + 
  scale_y_continuous(limits = c(-0.5, 0.5), 
                     expand = expand_scale(mult = c(0, 0))) + 
  scale_x_continuous(limits = c(-1.25, 1.25), 
                     expand = expand_scale(mult = c(0, 0))) + 
  theme(legend.position = "bottom", 
        legend.title = element_blank()) + 
  scale_fill_manual(values = c("orange", "red")) + 
  xlab("Dimension 1") + 
  ylab("Dimension 2")

yNorm3_cap <- "Multidimensional scaling (MDS) plot. Shows each LC-MS/MS
               biological replicate sample after normalisation, with the 
               first 2 dimensions plotted. Pairwise distance between samples 
               approximately indicates the log2 fold change between samples."
```

## Yeast EV versus WCL protein comparison

Lists of proteins quantified in at least 1/3 biological replicates of each 
condition are generated using:

```{r yVennData, echo=TRUE}
# filter for proteins quantified in min 1/3 reps of EV
y_ev <- filter_na(yeast_norm2, 
                  op = "<=", 
                  pat = "EV", 
                  val = 2)

# filter for proteins quantified in min 1/3 reps of WCL
y_wcl <- filter_na(yeast_norm2, 
                   op = "<=", 
                   pat = "W", 
                   val = 2)
```

These are then compared via a Venn diagram. As per section 2.4, every protein 
that appears in the Venn diagram has been quantified in at least 2/3 
replicates of EV or WCL.

```{r, echo=TRUE, eval=FALSE}
# define a comparison list for a Venn diagram
venn_comp <- list("EV" = rownames(y_ev), 
                  "WCL" = rownames(y_wcl))

# plot the Venn diagram
venn <- plot_venn(venn_comp, 
                  type = "plot")

grid::grid.draw(venn)
```

Of `r nrow(yeast_norm2)` proteins, 494 (51.7%) are quantified by at least 1/3 
biological replicates of both EV and WCL **and** at least 2/3 replicates of 
either EV or WCL. 

```{r yVenn}
# define comparison list for Venn diagram
venn1_comp <- list("EV" = rownames(y_ev), 
                   "WCL" = rownames(y_wcl))

# define ggplot2 default colour palette function
gg_colour <- function(n) {
  hues <- seq(15, 375, length = n + 1)
  grDevices::hcl(h = hues, l = 65, c = 100)[1:n]
}

# plot Venn diagram
venn1 <- plot_venn(venn1_comp, 
                   use_uniprot = FALSE, 
                   type = "plot", 
                   fontfamily = "sans", 
                   cat.fontfamily = "sans", 
                   cex = 1.9, 
                   cat.cex = 1.2, 
                   fill = gg_colour(2), 
                   alpha = c(rep(0.6, 2)))

# draw Venn diagram
grid::grid.draw(venn1)

venn_cap <- "Venn diagram of proteins quantified in each condition. 
             Proteins in this plot are quantified in at least 2/3 replicates 
             of either EV or WCL."
```

Next, the most abundant proteins that are only detected in DAY286 yeast 
EVs and not in the WCL can be identified.

```{r yTop10Data}
# get top 10 most abundant protein exclusive to yeast EVs
# in descending order of abundance
yeast_top10 <- y_ev %>% 
  tibble::as_tibble(rownames = "id") %>% 
  filter_na(., op = "==", pat = "W", val = 3) %>% 
  select(-contains("W")) %>% 
  mutate(mean = rowMeans(select(., contains("EV")), na.rm = TRUE)) %>% 
  top_n(n = 10, wt = mean) %>% 
  arrange(desc(mean)) %>% 
  match_id_multi(., "id", uniprot, "UP_accession", 
                 new = c("CGD_gene_name", 
                         "CGD_description")) %>% 
  select(CGD_gene_name, CGD_description, everything(), -id)
```

```{r yTop10}
knitr::kable(yeast_top10, 
             caption = "Top 10 most abundant DAY286 yeast EV proteins. They 
                        are not detected in DAY286 yeast WCL.", 
             col.names = c("Name", "Function", "EV2 LFQ", 
                           "EV3 LFQ", "EV4 LFQ", "Mean LFQ"), 
             digits = 1)
```

## Imputation of missing values

Missing values in the data will be imputed prior to the 
differential abundance analysis. Before imputation the percentage of missing 
values in each sample is explored again with `sum_table(yeast_norm2)`.

```{r ySum}
# explore percentage missing values in each sample
knitr::kable(sum_table(yeast_norm2), 
             caption = "Descriptive statistics for each LC-MS/MS sample.", 
             col.names = c("Sample", "Mean LFQ intensity", 
                           "Median LFQ intensity", "SD", "CV", 
                           "Min LFQ intensity", "Max LFQ intensity", 
                           "Range", "Number of proteins", "Number of NA", 
                           "Percent NA"), 
             digits = 1)
```

The percentage of missing values ranges from 21.9% up to 49.4% which is 
quite high, typical of label-free quantitative proteomics data[^3]. The 
pattern of missing values across the data is visualised using:

[^3]: Lazar, C., Gatto, L., Ferro, M., Bruley, C., Burger, T. (2016). 
Accounting for the Multiple Natures of Missing Values in Label-Free 
Quantitative Proteomics Data Sets to Compare Imputation Strategies. 
_Journal of Proteome Research, 15_(4), 1116-1125. 
[doi:10.1021/acs.jproteome.5b00981](https://doi.org/10.1021/acs.jproteome.5b00981)

```{r, echo=TRUE, eval=FALSE}
# visualise missing values using a heatmap
plot_missval2(yeast_norm2)
```

```{r yMissVal, fig.cap=missval_cap, fig.height=8}
plot_missval2(yeast_norm2)

missval_cap <- "Pattern of missing LFQ intensity values."
```

There are large chunks of data missing in either all of 
the WCL samples or all of the EV samples. To minimise the percentage of 
missing values in each sample, these proteins are removed prior to 
imputation and hence are not imputed.

```{r yExcl, echo=TRUE}
# should not impute with so many missing values
# select proteins with 3/3 missing values in EV or WCL
# these will not be imputed
yeast_excl <- filter_na2(yeast_norm2, 
                         logic = "or", 
                         op = "==", 
                         pat1 = "EV", val1 = 3, 
                         pat2 = "W", val2 = 3)
```

Only proteins quantified in at least 1/3 replicates of both EV and WCL will be 
retained for imputation.

```{r yBoth, echo=TRUE}
# filter for proteins with at least 1 valid value in EV and WCL
yeast_both <- filter_na2(yeast_norm2, 
                         logic = "and", 
                         op = "<=", 
                         pat1 = "EV", val1 = 2, 
                         pat2 = "W", val2 = 2)
```

`r nrow(yeast_excl)` proteins which have 3/3 missing values in EV or 
WCL are removed (`yeast_excl`). 
Conversely, `r nrow(yeast_both)` proteins are kept to have missing values 
imputed (`yeast_both`). The percentage of missing values across each sample 
in `yeast_both` now ranges from 2.2% to 31.4%. 

```{r ySum2}
knitr::kable(sum_table(yeast_both), 
             caption = "Descriptive statistics for each LC-MS/MS sample.", 
             col.names = c("Sample", "Mean LFQ intensity", 
                           "Median LFQ intensity", "SD", "CV", 
                           "Min LFQ intensity", "Max LFQ intensity", 
                           "Range", "Number of proteins", "Number of NA", 
                           "Percent NA"), 
             digits = 1)
```

The pattern of missing values in `yeast_both` seems to be biased to certain 
samples, especially WCL replicate 2. Therefore, 
these proteins may be missing not at random (MNAR). 

```{r yMissVal2, fig.cap=missval2_cap , fig.height=8}
plot_missval2(yeast_both)

missval2_cap <- "Pattern of missing LFQ intensity values."
```

The LFQ intensity distributions and cumulative fraction of proteins with and 
without missing values can be plotted with: 

```{r, echo=TRUE, eval=FALSE}
# detect intensity differences for proteins with missing values
plot_detect2(yeast_both)
```

```{r bSum3, fig.cap=detect_cap, fig.wide=TRUE}
plot_detect2(yeast_both)

detect_cap <- "LFQ intensities of proteins with and without missing values."
```

These plots show that missing values are biased to proteins with lower 
average intensities, and are likely MNAR. Hence, a left-censored imputation 
method is used for imputation of missing values. 

```{r, echo=TRUE, eval=FALSE}
# impute missing values with QRILC method
yeast_imp <- impute_QRILC(yeast_both)
```

```{r, echo=TRUE, eval=FALSE}
plot_imputation2(yeast_exp, 
                 facet_labels = c(yeast_both = "Before imputation", 
                                  yeast_imp = "After imputation"), 
                 yeast_both, 
                 yeast_imp)
```

```{r yImpute, fig.cap=imp_cap, fig.wide=TRUE}
# proteins with missing values tend to have lower intensity
# therefore proteins are MNAR, close to detection limit

# use left censored imputation method
yeast_imp <- impute_QRILC(yeast_both)

plot_imputation2(yeast_exp, 
                 facet_labels = c(yeast_both = "Before imputation", 
                                  yeast_imp = "After imputation"), 
                 yeast_both, yeast_imp)

imp_cap <- "LFQ intensity distribution before and after imputation."
```

## Differential abundance analysis with _limma_

```{r yeastDE}
# recombine imputed proteins and non-imputed proteins in matrix
yeast_de <- rbind(yeast_excl, yeast_imp)

# see limma user guide section 9.2 for more info about DE
# create design matrix
y_samp <- data.frame(T = (rep(c("EV", "WCL"), each = 3)))

y_design <- stats::model.matrix(~ 0 + T, data = y_samp)
colnames(y_design) <- c("EV", "WCL")

y_contrasts <- c("EV - WCL")

# make all pair-wise comparisons for specified contrasts
# and perform limma::eBayes()
y_efit <- limma_eBayes(yeast_de, y_design, y_contrasts)

# extract DE results 
yeast_res <- get_results(efit = y_efit, 
                         mat = yeast_de, 
                         p_val = 0.01, 
                         lfc = 0, 
                         type = "individual")[[1]]

yeast_res_short <- yeast_res %>% 
  select(CGD_gene_name, Protein_name, logFC, contains("_mean"), adj.P.Val, group) %>% 
  mutate_at(.vars = vars(logFC, contains("_mean")), 
            .funs = list(~round(., digits = 2))) %>% 
  mutate(adj.P.Val = signif(adj.P.Val, digits = 3), 
         group = as.factor(group))

DT::datatable(yeast_res_short, 
              caption = "This is a table caption. Note that adjusted p values have been shortened
              to three significant figures for ease of display.", 
              rownames = FALSE, 
              colnames = c("Gene" = "CGD_gene_name", 
                           "Protein" = "Protein_name", 
                           "p.adj" = "adj.P.Val"), 
              filter = list(position = "bottom", clear = FALSE, plain = TRUE), 
              class = "display compact", 
              options = list(
                autowidth = TRUE, 
                columnDefs = list(list(width = "40%", targets = 1), 
                                  list(width = "10%", targets = c(0, 2:6)), 
                                  list(targets = 1, 
                                       render = DT::JS(
                                         "function(data, type, row, meta) {",
                                         "return type === 'display' && data.length > 40 ?",
                                         "'<span title=\"' + data + '\">' + data.substr(0, 40) + '...</span>' : data;",
                                         "}")), 
                                  list(targets = 0, 
                                       render = DT::JS("function(data, type, row, meta) {",
                                         "return type === 'display' && data.length > 12 ?",
                                         "'<span title=\"' + data + '\">' + data.substr(0, 12) + '...</span>' : data;",
                                         "}"))
                                  )
                )
              )


```




# Comparison to previous data


TEXT

```{r yeastComp}
g_list <- gilbona %>% 
  filter(EV1_peptides > 0 | EV2_peptides > 0 | EV3_peptides > 0) %>% 
  mutate(gn = match_id(UP_accession, uniprot, "UP_accession", "CGD_gene_name")) %>% 
  pull(gn)

v_list <- vargas %>% 
  filter(S11_peptides > 0) %>% 
  mutate(gn = match_id(UP_accession, uniprot, "UP_accession", "CGD_gene_name")) %>% 
  pull(gn)

w_list <- wolf %>% 
  filter(WT_pep > 0) %>% 
  mutate(gn = match_id(UP_accession, uniprot, "UP_accession", "CGD_gene_name")) %>% 
  pull(gn)

z_list_y <- zarnowski %>% 
  filter(Planktonic_NTS > 0) %>% 
  mutate(gn = match_id(UP_accession, uniprot, "UP_accession", "CGD_gene_name")) %>% 
  pull(gn)

y_ev_list <- match_id(rownames(y_ev), uniprot, "UP_accession", "CGD_gene_name", concat = FALSE)

yeast_comp <- list("This paper" = y_ev_list, "Gil-Bona" = g_list, "Vargas" = v_list, "Wolf" = w_list, "Zarnowski" = z_list_y)

upset(fromList(yeast_comp), 
      point.size = 4.5, 
      line.size = 2, 
      mainbar.y.label = "\n\nIntersection protein number", 
      mainbar.y.max = 400, 
      sets.x.label = "Total protein number", 
      text.scale = c(2, 2, 1.2, 1.2, 1.7, 2), 
      order.by = "degree", 
      queries = list(list(query = intersects, 
                          params = list(c("This paper", "Gil-Bona", "Vargas", "Wolf", "Zarnowski")), 
                          color = "limegreen", 
                          active = TRUE)))
yeast_comp_tab <- plot_venn(vlist = yeast_comp, use_uniprot = FALSE, type = "list") %>% 
  mutate(..values.. = lapply(..values.., function(x) paste(x, collapse = ", ")))

knitr::kable(yeast_comp_tab[1,], 
             caption = "A table caption")



```



















