---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo=FALSE}
knitr::opts_chunk$set(collapse = TRUE,
                      comment = "#>")
```

# R

## Description

This directory contains R script files used throughout the project. Each file 
should only contain 1 function or dataset. Many of these functions have been
taken from the _Bioconductor_ package 
**[DEP](https://doi.org/10.18129/B9.bioc.DEP)**^1^, and adapted for use on
basic matrices instead of `SummarizedExperiment` objects.

After installing `candidaev`, an index of all the included 
functions/datasets can be viewed in R using:

```{r, eval=FALSE}
help(package = "candidaev")
```

### References

1. Zhang, X., Smits, A., van Tilburg, G., Ovaa, H., Huber, W., Vermeulen, 
   M. (2018). Proteome-wide identification of ubiquitin interactions using UbIA-MS. 
   _Nature Protocols_, **13**, 530-550. 
   [doi:10.1038/nprot.2017.147](https://doi.org/10.1038/nprot.2017.147)
   
